On Math

### 2025-04-25 cards/paper.tex
updates

\begin{equation}
  \frac{a}{\sin A} = \frac{b}{\sin B} = \frac{c}{\sin C}
\end{equation}

and also:

\begin{equation}
  \frac{\sin A}{a} = \frac{\sin B}{b} = \frac{\sin C}{c}
\end{equation}

When we have obtuse angle we need to subtract the result from $180$.

\section{Complex numbers}
Complex numbers can be represented by the point $(x, y)$ where $x$ is real axis and $y$
is imaginary axis. We call this plane as a 'complex plane', also called 'Argand diagram',
denoted as $C$.

@@

\subsection{Argument of a complex number}
There is natural angle $\theta$ that \textit{z} makes with real x-axis. This angle is called \textit{the argument of z}. The numeric value of $arg(z)$ is given in radians and also
\begin{equation}
0 \leq arg(z) \leq 2\pi
\end{equation}

Example:
\begin{equation}
  \begin{gathered}
  z = 3 + 4i \\
  \tan \theta = \frac{4}{3} \\
  \theta = \arctan\frac{4}{3} \approx 0.93
  \end{gathered}
\end{equation}

\section{Geometry}
\subsection{Rigid motions}
Or sometimes - any transformations that does not change the distances
and angles between points in a figure. They are ANGLE-PRESERVING and
DISTANCE-PRESERVING.

Rigid motions types:
- translations
- rotations
- reflections

Not a rigid motions:
- dilations
- stretches

@@

$g$ - $9.8\frac{m}{s^2}$ - constant acceleration due to gravity
\section{Functions}
\subsection{Quadratic formula}
\begin{equation}
  x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

\subsection{Odd and even functions}

Function is said to be odd if:
\begin{equation}
  f(-x) = -f(x)
\end{equation}

Function is said to be even if:
\begin{equation}
  f(-x) = f(x)
\end{equation}
\subsection{Factor cubics by grouping}

$3x^3-6x^2+2x-4$

1. group them

$(3x^3-6x^2)+(2x-4)$

2. factor larger terms so the leftover match

$3x^2(x-2)+2(x-2)$

3. factor out the binomial

$(3x^2+2)(x-2)$

Note: not all cubics can be factored, however,
of the cubics that can be factored,
many can be factored by grouping

\subsection{Factor theorem}
For polynomial, if we know root, we can determine a factor of the
polynomial. If $p(x)$ is a polynomial and $x=r$ is a root of $p(x)$ then
\begin{equation}
  \begin{gathered}
  p(r) = 0 \\
  so (x-r) is a factor of p(x) \\
  generally: (ax-b) is a factor of a polynomial p(x) then \\
  x = \frac{b}{a} is a root of p(x)
\end{gathered}
\end{equation}

\subsection{Rational roots theorem}
Support the rational number $\frac{p}{q}$ in lowest terms, is a root of a polynomial
with integer coefficient.
% TODO: add colors to differentiate
Then 'p' must be the factor of the polynomial's constant term, and 'q' must be a factor
of its leading coefficient.
Given a polynomial with integer coefficient, any rational roots of the polynomial must take the
form:
\begin{equation}
  \pm \frac{factor of the constant term}{factor of the leading coefficient}
\end{equation}

For example:
\begin{equation}
  f(x) = 8x^3 + 12x^2 + 6x + 2
\end{equation}
so we have 16 possible roots:
\begin{equation}
  \begin{gathered}
    % TODO: strike through
    \pm \frac{1}{1}, \pm \frac{1}{2}, \pm \frac{1}{4}, \pm \frac{1}{8}, \\
    \pm \frac{2}{1}, \pm \frac{2}{2}, \pm \frac{2}{4}, \pm \frac{2}{8}
  \end{gathered}
\end{equation}

Removing the duplicates, we have 10 possible rational roots:
\begin{equation}
  \pm \frac{1}{8}, \pm \frac{1}{4}, \pm \frac{1}{2}, \pm 1, \pm 2
\end{equation}

Note: it is about 'possible' rational roots.

\subsection{Function composition}
\begin{equation}
  (f \circ g)(x) = f(g(x))
\end{equation}
\subsection{Inverse functions}
Inverse functions are denoted by $f^{-1}(x)$.

For $f(x)$ that maps $x=2$ to $y=0$ an inverse function $f^{-1}(x)$ instantiation would be
$f^{-1}(0)=0$.

Generally it is a simple composition:
\begin{equation}
  \begin{gathered}
    f^{-1}(f(x)) = x \\
    so \\
    (f^{-1} \circ f)(x) = x
  \end{gathered}
\end{equation}

Calculate inverse function
1. swap $x$ and $y$
2. solve for $y$

@@

Domain and range of inverse function:
Domain of a function $f(x)$ is the range of its inverse function $f^{-1}(x)$.
Range of a function $f(x)$ is the domain of its inverse function $f^{-1}(x)$.

\subsection{Multiplicities of the roots of a polynomial}
\begin{equation}
  p(x) = (x-5)^2(x+8)
\end{equation}
for that polynomial roots are $x=5$ and $x=-8$ but factor $(x-5)^2$
appears TWICE so it is a MULTIPLE ROOT
$(x+8)$ on the other hand is a SIMPLE ROOT.

@@

\subsection{Rational functions}
Horizontal asymptotes
1. Identify the 'dominant term' which is a leading term of the polynomial though we
usually ugnore the coefficients.

@@

\subsection{Derivative of the (natural) logarithm}
For natural logarithm:
\begin{equation}
  \frac{d}{dx}(\ln x) = \frac{1}{x}
\end{equation}

For general logarithm:
\begin{equation}
  \frac{d}{dx}(\log_a x) = \frac{1}{x\ln a}
\end{equation}

Example with changing base:
\begin{equation}
  \begin{gathered}
  \log_ax = \frac{\ln x}{\ln a} \\
  \frac{d}{dx}(\log_ax) = \frac{d}{dx}(\frac{\ln x}{\ln a}) = \\
  = \frac{1}{\ln a} * \frac{d}{dx}(\ln x) = \\
  = \frac{1}{\ln a} * \frac{1}{x} = \frac{1}{x\ln a}
  \end{gathered}
\end{equation}
\subsection{Chain rule for logarithmic function}
\begin{equation}
  \begin{gathered}
  y = \ln(2x+1) \\
  y = \ln u \\
  u = 2x+1 \\
  y' = \frac{1}{u} \\
  u' = 2 \\
  \frac{dy}{dx} = \frac{1}{u} * 2 = \frac{2}{u} = \frac{2}{2x+1}
  \end{gathered}
\end{equation}

>>>>

\begin{equation}
  \frac{a}{\sin A} = \frac{b}{\sin B} = \frac{c}{\sin C}
\end{equation}
and also:
\begin{equation}
  \frac{\sin A}{a} = \frac{\sin B}{b} = \frac{\sin C}{c}
\end{equation}
When we have obtuse angle we need to subtract the result from $180$.

\section{Complex numbers}
Complex numbers can be represented by the point $(x, y)$ where $x$ is real axis and $y$
is imaginary axis. We call this plane as a \textbf{complex plane}, also called
\textit{Argand diagram}, denoted as $C$.

@@

\subsection{Argument of a complex number}
There is natural angle $\theta$ that \textit{z} makes with real x-axis. This angle is called \textit{the argument of z}. The numeric value of $arg(z)$ is given in radians and also
\begin{equation}
0 \leq arg(z) \leq 2\pi
\end{equation}
Example:
\begin{equation}
  \begin{gathered}
  z = 3 + 4i \\
  \tan \theta = \frac{4}{3} \\
  \theta = \arctan\frac{4}{3} \approx 0.93
  \end{gathered}
\end{equation}

\section{Geometry}
\subsection{Rigid motions}
Or sometimes - any transformations that does not change the distances
and angles between points in a figure. They are \textbf{angle-preserving} and
\textbf{distance-preserving}.

Rigid motions types:

- translations

- rotations

- reflections

Not a rigid motions:

- dilations

- stretches

@@

$g$ - $9.8\frac{m}{s^2}$ - constant acceleration due to gravity
\section{Functions}
\subsection{Standard form}
\begin{equation}
  ax + by = c
\end{equation}
\subsection{Parallel lines}
Calculate equation of the parallel line:
\begin{equation}
  y - y_1 = m(x - x_1)
\end{equation}
Calculate a slope of a line given two points:
\begin{equation}
  m = \frac{y_2 - y_1}{x_2 - x_1}
\end{equation}
\subsection{Quadratic formula}
\begin{equation}
  x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

\subsection{Odd and even functions}
Function is said to be odd if:
\begin{equation}
  f(-x) = -f(x)
\end{equation}
Function is said to be even if:
\begin{equation}
  f(-x) = f(x)
\end{equation}
\subsection{Factor cubics by grouping}
For example to factor following function by grouping:

\begin{equation}
  3x^3-6x^2+2x-4
\end{equation}
1. group them
\begin{equation}
  (3x^3-6x^2)+(2x-4)
\end{equation}
2. factor larger terms so the leftover match
\begin{equation}
  3x^2(x-2)+2(x-2)
\end{equation}
3. factor out the binomial
\begin{equation}
  (3x^2+2)(x-2)
\end{equation}
Note: not all cubics can be factored, however,
of the cubics that can be factored,
many can be factored by grouping

\subsection{Factor theorem}
For polynomial, if we know root, we can determine a factor of the
polynomial. If $p(x)$ is a polynomial and $x=r$ is a root of $p(x)$ then
\begin{equation}
  p(r) = 0
\end{equation}
so $(x-r)$ is a factor of $p(x)$
generally: $(ax-b)$ is a factor of a polynomial $p(x)$ then:
\begin{equation}
  x = \frac{b}{a} is a root of p(x)
\end{equation}

\subsection{Rational roots theorem}
Support the rational number $\frac{p}{q}$ in lowest terms, is a root of a polynomial
with integer coefficient.
% TODO: add colors to differentiate
Then 'p' must be the factor of the polynomial's constant term, and 'q' must be a factor
of its leading coefficient.
Given a polynomial with integer coefficient, any rational roots of the polynomial must take the
form:
\begin{equation}
  \pm \frac{factor  of  the  constant  term}{factor  of  the  leading  coefficient}
\end{equation}
For example:
\begin{equation}
  f(x) = 8x^3 + 12x^2 + 6x + 2
\end{equation}
so we have 16 possible roots:
\begin{equation}
  \begin{gathered}
    % TODO: strike through
    \pm \frac{1}{1}, \pm \frac{1}{2}, \pm \frac{1}{4}, \pm \frac{1}{8}, \\
    \pm \frac{2}{1}, \pm \frac{2}{2}, \pm \frac{2}{4}, \pm \frac{2}{8}
  \end{gathered}
\end{equation}
Removing the duplicates, we have 10 possible rational roots:
\begin{equation}
  \pm \frac{1}{8}, \pm \frac{1}{4}, \pm \frac{1}{2}, \pm 1, \pm 2
\end{equation}
Note: it is about \textbf{possible} rational roots.

\subsection{Function composition}
\begin{equation}
  (f \circ g)(x) = f(g(x))
\end{equation}
\subsection{Inverse functions}
Inverse functions are denoted by $f^{-1}(x)$.
For $f(x)$ that maps $x=2$ to $y=0$ an inverse function $f^{-1}(x)$ instantiation would be
$f^{-1}(0)=0$.
Generally it is a simple composition:
\begin{equation}
  f^{-1}(f(x)) = x
\end{equation}
so
\begin{equation}
  (f^{-1} \circ f)(x) = x
\end{equation}

Calculate inverse function:
1. swap $x$ and $y$

2. solve for $y$

@@

Domain of a function $f(x)$ is the range of its inverse function $f^{-1}(x)$.

Range of a function $f(x)$ is the domain of its inverse function $f^{-1}(x)$.

\subsection{Multiplicities of the roots of a polynomial}
\begin{equation}
  p(x) = (x-5)^2(x+8)
\end{equation}
for that polynomial roots are $x=5$ and $x=-8$ but factor $(x-5)^2$
appears \textbf{twice} so it is a \textbf{multiple root}
$(x+8)$ on the other hand is a \textbf{simple root}.

@@

\subsection{Rational functions}
Horizontal asymptotes

1. Identify the 'dominant term' which is a leading term of the polynomial though we
usually ugnore the coefficients.

@@

\subsection{Derivative of the (natural) logarithm}
For natural logarithm:
\begin{equation}
  \frac{d}{dx}(\ln x) = \frac{1}{x}
\end{equation}
For general logarithm:
\begin{equation}
  \frac{d}{dx}(\log_a x) = \frac{1}{x\ln a}
\end{equation}
Example with changing base:
\begin{equation}
  \begin{gathered}
  \log_ax = \frac{\ln x}{\ln a} \\
  \frac{d}{dx}(\log_ax) = \frac{d}{dx}(\frac{\ln x}{\ln a}) = \\
  = \frac{1}{\ln a} * \frac{d}{dx}(\ln x) = \\
  = \frac{1}{\ln a} * \frac{1}{x} = \frac{1}{x\ln a}
  \end{gathered}
\end{equation}
\subsection{Chain rule for logarithmic function}
\begin{equation}
  \begin{gathered}
  y = \ln(2x+1) \\
  y = \ln u \\
  u = 2x+1 \\
  y' = \frac{1}{u} \\
  u' = 2 \\
  \frac{dy}{dx} = \frac{1}{u} * 2 = \frac{2}{u} = \frac{2}{2x+1}
  \end{gathered}
\end{equation}
